## Modelling Viral Load using TREIV-PKPD Model
## Create Fig 2, S-Fig 2, S-Fig 3

```{r}
library(dplyr)
library(deSolve)
library(ggplot2)
library(MCMCglmm)
library(MASS)
```

######### Set the time range #########
```{r}
Tmin <- 0
Tmax <- 40
step_size <- 0.1
times <- seq(Tmin, Tmax, step_size)
```

########## Read in parameters #############
```{r}
pop = read.csv("populationParameters.txt")
pop
```
######### Sample function to calculate CI ##############

```{r}
sample_pars_pop <- function(pop, num){
  
  logit <- function(p) log(p / (1 - p))
  inv_logit <- function(x) exp(x) / (1 + exp(x))
 
  # Identify rows with log10
  log10_pop <- grep("^log10", pop$parameter)        # log10 in _pop
  log10_omega <- grep("omega_log10", pop$parameter)       # omega_log10
  
  # Identify rows without log10
  non_log10_pop <- setdiff(grep("_pop$", pop$parameter), log10_pop)      # non-log10 _pop
  non_log10_omega <- setdiff(grep("^omega", pop$parameter), log10_omega)

  n_delta <- non_log10_pop[1]
  omega_delta <- non_log10_omega[1]

  n_tau <- non_log10_pop[2]
  omega_tau <- non_log10_omega[2]
 
  n_prf <- non_log10_pop[3]
  omega_prf <- non_log10_omega[3]
  
  pars <- matrix(0, num, length(log10_pop)+length(non_log10_pop))
  
  for (i in 1:(length(log10_pop)+length(non_log10_pop))) {
    
    if (i <= length(log10_pop)){
      
      mean_par <- pop$value[log10_pop[i]]
      sd_par <- pop$value[log10_omega[i]]
      lower <- mean_par-1.96*sd_par
      upper <- mean_par+1.96*sd_par
      pars[, i] <- 10^(rtnorm(n=num, mean=mean_par, sd=sd_par, lower=lower, upper=upper))
   
    } else if (i == length(log10_pop)+1 & length(log10_pop)+1 == n_delta ) {  # Sample for delta 
      
      mean_par <- pop$value[non_log10_pop[i-length(log10_pop)]]
      sd_par <- pop$value[non_log10_omega[i-length(log10_pop)]]
      lower <- mean_par-1.96*sd_par
      upper <- mean_par+1.96*sd_par
      pars[, i] <- exp(rtnorm(n=num, mean=log(mean_par), sd=sd_par, lower=log(lower), upper=log(upper)))
    
    } else if (i == length(log10_pop)+2 & length(log10_pop)+2 == n_tau) {  # Sample for tau
      logit <- function(p) log(p / (20 - p))
      inv_logit <- function(x) 20*exp(x) / (1 + exp(x))
      mean_par <- pop$value[non_log10_pop[i-length(log10_pop)]]
      sd_par <- pop$value[non_log10_omega[i-length(log10_pop)]]
      pars[, i] <- (inv_logit(rnorm(n=num, mean=logit(mean_par), sd=sd_par)))  # incubation period ranges from 0 to 20
      
    } else if (i == length(log10_pop)+3 & length(log10_pop)+3 == n_prf ) {  # Sample for prf
      
      mean_par <- pop$value[non_log10_pop[i-length(log10_pop)]]
      sd_par <- pop$value[non_log10_omega[i-length(log10_pop)]]
      lower <- mean_par-1.96*sd_par
      upper <- mean_par+1.96*sd_par
      pars[, i] <- exp(rtnorm(n=num, mean=log(mean_par), sd=sd_par, lower=log(lower), upper=log(upper)))
    }
  }    
  return(pars)

}    
```

########### TREIV-PKPD FUNCTION #############
```{r}

Intfun_ept_hill <- function(pars, t_treat, antiviral) {
  
  beta <- as.numeric(pars[1])
  pi <- as.numeric(pars[2])
  phi <- as.numeric(pars[3])
  rho <- as.numeric(pars[4])
  delta <- as.numeric(pars[5])
   # To be estimated
  tau <- round(as.numeric(pars[6]),digit=0)    # round the tau, as the same digit of stepsize
  #tau <- 1 #population one
  prf <- as.numeric(pars[7])
  
  A <- round(antiviral, digit = 0)
  # Fixed PKPD parameters from the paper
   
  E_max <- 0.999
  K_e <- 9.98
  K_PL <- 1.58
  K_LP <- 1.22
  K_CL <- 4.96
  Vol <- 41743
  MolMass <- 499.5
  n <- 3.16
  IC50 <- 0.083  # in µM (43.6 nM converted)
  Dose <- 300     # in mg
  
  # Time settings
  t_treat <- round(t_treat, digit =1 )
  rec <- t_treat + tau	  # Population taa fixed
  Tmin <- 0
  Tmax <- 40
  step_size <- 0.1
  times <- seq(Tmin, Tmax, step_size)
  
  # Treatment indicator (Antiviral == 1 during treatment window)
  treatment <- data.frame(times = times, h = rep(0, length(times)))
  treatment$h <- ifelse(times >= rec & times < rec + 5, A, 0)  # 5-day treatment
  h_t2 <- approxfun(treatment$times, treatment$h, rule = 2)
  
  # Dosing events: 300 mg every 12 hours (0.5 days) for 5 days (10 doses)
  dose_times <- seq(rec, rec + 4.5, by = 0.5)  # t_treat, t_treat+0.5, ..., t_treat+4.5
  dose_events <- data.frame(var = "A_G", time = dose_times, value = Dose*A, method = "add")
  #dose_events$time <- round(dose_events$time, digit = 1)
  # ODE system
  derivs <- function(times, y, pars, h_t2) {
    with(as.list(c(pars, y)), {
      # Treatment indicator
      h2 <- h_t2(times)
      
      # PK equations
      dA_G <- -K_e * A_G
      dA_P <- K_e * A_G + K_LP * A_L - (K_CL + K_PL) * A_P
      dA_L <- K_PL * A_P - K_LP * A_L
      
      # Compute plasma concentration
      C_P_mg_per_mL <- A_P / Vol
      C_P_uM <- C_P_mg_per_mL * (1e6 / MolMass)
      
      # Compute efficacy (eps)
      eps <- E_max * (C_P_uM^n) / ((prf * IC50)^n + C_P_uM^n)
      
      # Apply treatment window (eps_t)
      eps_t <- ifelse(h2 == 1, eps, 0)
      
      # Viral dynamics equations
      dT <- -beta * T * V - phi * I * T + rho * R
      dR <- phi * I * T - rho * R
      dE <- beta * T * V - 4 * E
      dI <- 4 * E - delta * I
      dV <- (1 - eps_t) * pi * I - 15 * V
      
      return(list(c(dT, dR, dE, dI, dV, dA_G, dA_P, dA_L), eps_t = eps_t))
    })
  }
  
  # Initial conditions
  y <- c(T = 10^7, R = 0, E = 0, I = 0, V = 1, A_G = 0, A_P = 0, A_L = 0)
  
  out <- ode(y = y, times = round(times,digit=1), func = derivs, parms = pars, h_t2 = h_t2, events = list(data = dose_events), method = "bdf")
  
  out_df <- as.data.frame(cbind(time = out[,1], aV1 = log10(out[,6]), eps_t = out[,"eps_t"], Ap = out[,8]))
  
  return(out_df)
}
```

################ Function to compute average efficacy #####################

```{r}
#average efficacy
ComputeEave <- function( out, t) {  #t = time of treatment since INFECTION
                                                  # out: dataframe contains time, and eps_t
  
  #tau <- round(as.numeric(pars[6]), digit =1 )
  # Treatment period
  t_start <- t
  t_end <- t_start + 5  # 5-day treatment
  
  treatment_period <- out[out$time >= t_start & out$time <= t_end, ]
  
 # if (nrow(treatment_period) < 2) {
 #   stop("Not enough data points in the treatment period to compute E_ave.")
 # }
  
  times <- round(treatment_period$time,digit =1)
  eps <- treatment_period$eps_t
  
  integral <- 0
  for (i in 1:(length(times) - 1)) {
    dt <- times[i + 1] - times[i]
    integral <- integral + 0.5 * (eps[i] + eps[i + 1]) * dt
  }
  
  E_ave <- integral / (t_end - t_start)
  
  return(E_ave)
}
```

######## Function to store data in column format ############
```{r}
run_ODE_control <- function(pars,t,antiviral){
  time <- t
  A <- antiviral
  total_VL <- matrix(NA,nrow=(length(seq(Tmin,Tmax,step_size))),ncol=num)
  for(i in 1:num){
    out <- Intfun_ept_hill(as.numeric(pars[i,]), t_treat = time, A)
  total_VL[,i] <- out$aV1
  }
  return(total_VL)
}
```

######## Set Treatment Conditions for Demonstration ##################
```{r}
df_cov <- data.frame( group= c("Treatment Date: 0",   # 0 days after dection
                               "Treatment Date: 5",
                               "Treatment Date: 0",  
                               "Treatment Date: 5" 
                              ),    
                 t_treat = c(0,5,0,5),
                 antiviral = c (1,1,0,0))
df_cov 
```

############ Run Function ####################
```{r}
num = 1000 #1000 simulations
Tmin <- 0
Tmax <- 40
step_size <- 0.1
times<-c(seq(Tmin,Tmax,step_size))

#inv_logit <- function(x) exp(x) / (1 + exp(x))
#inv_logit_20 <- function(x) 20*exp(x) / (1 + exp(x))

pb <- txtProgressBar(min = 0, max = 4, style = 3)

Fit <- list()


for (g in 1:4){
  
  group <- df_cov$group[g]
  t_treat <- df_cov$t_treat[g]
  antiviral <- df_cov$antiviral[g]
  
  pars <- c(beta=10^pop$value[1],  
           pi=10^pop$value[2],
           phi=10^pop$value[3],
           rho=10^pop$value[4],
           delta=pop$value[5],
           tau=pop$value[6],
           prf=pop$value[7])
  
  t_treat_since_inf <- t_treat +  round(pop$value[6], digit =0 )
  
  fit <- Intfun_ept_hill(pars, t = t_treat, antiviral)

  best_fit <- fit[,c(1,2)]
  efficacy <- fit[,3]
  Drug_conc <- fit[,4]
  pars <- sample_pars_pop(pop, num)
  total_VL <- run_ODE_control(pars,t_treat, antiviral)
  
  Min  <- apply(total_VL,1,function(x){quantile(x,0.025,na.rm=T)})
  Max  <- apply(total_VL,1,function(x){quantile(x,0.975,na.rm=T)})
  Min50  <- apply(total_VL,1,function(x){quantile(x,0.25,na.rm=T)})
  Max50  <- apply(total_VL,1,function(x){quantile(x,0.75,na.rm=T)})
  Mean50  <- apply(total_VL,1,function(x){quantile(x,0.50,na.rm=T)})
  index <- which(Mean50 < -2)[1]
  # Check if an index was found, and then replace the value
  if (!is.na(index)) {
      Mean50[index] <- -2
    }  
  
  Fit[[g]] <- cbind(best_fit,efficacy,Drug_conc,Min,Max,Min50,Max50,Mean50,group,t_treat_since_inf, antiviral)
  Fit[[g]] <- data.frame(Fit[[g]])
  colnames(Fit[[g]]) <- c("time","best_fit","eps_t","Drug_conc.","Min95","Max95","Min50","Max50","Mean50","t_treat","t_treat_since_inf", "antiviral")
  Fit[[g]]$Max95[Fit[[g]]$Max95 > 11] <- 11
  Fit[[g]]$Max50[Fit[[g]]$Max50 > 11] <- 11
  Fit[[g]]$Min95[Fit[[g]]$Min95 < -2] <- -2
  Fit[[g]]$Min50[Fit[[g]]$Min50 < -2] <- -2

  setTxtProgressBar(pb, g)
}
close(pb)

combine_pop <- map_df(Fit, ~as.data.frame(.x))

```

######################################
##########Figure 3####################
######################################
```{r}
plot_pop <- ggplot(combine_pop, aes(x = time, y = best_fit, group = antiviral)) +
  geom_line(aes(color = antiviral), size = 0.75, alpha = 0.8) +

  geom_ribbon(data = subset(combine_pop, antiviral == "1"),
              aes(ymin = Min95, ymax = Max95, fill = antiviral), alpha = 0.1) +
  geom_ribbon(data = subset(combine_pop, antiviral == "1"),
              aes(ymin = Min50, ymax = Max50, fill = antiviral), alpha = 0.2) +

  # Vertical lines
  geom_vline(aes(xintercept = t_treat_since_inf), color = "green", linetype = "dotted", size = 1.2) +
  geom_vline(aes(xintercept = t_treat_since_inf + 5), color = "green", linetype = "dotted", size = 1.2) +
  geom_hline(aes(yintercept = 1.31), color = "grey", linetype = "dotted", size = 1.2) +

  # Facet by t_treat
  facet_wrap(~ t_treat, scales = "fixed") +

  scale_color_manual(values = c("1" = "#d62728", "0" = "#1f77b4"),  # treated = red, untreated = blue
                     labels = c("0" = "Untreated", "1" = "Treated")) +
  scale_fill_manual(values = c("1" = "#d62728", "0" = "#1f77b4"),
                    labels = c("0" = "Untreated", "1" = "Treated")) +

  # Labels and theme
  labs(
    x = "Days after Infection",
    y = "log10 Viral Load (RNA copies/ml)",
    color = "Antiviral"#, fill = "Antiviral"
  ) +
 theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='none',
        #axis.title.y = element_text(size=11,family="sans"),
       # axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  )+ theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 14),        # X-axis label font size
  axis.title.y = element_text(size = 14),        # Y-axis label font size
  axis.text = element_text(size = 14) ,  
  strip.text = element_text(size = 14),# Axis tick labels font size
  legend.title = element_text(size = 14),  # Legend title font size
  legend.text = element_text(size = 12),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
) + ylim(-2,11) + xlim(0,30)

plot_pop
```

```{r}
plot_pop_drug <- ggplot(combine_pop, aes(x = time, y = Drug_conc., group = antiviral)) +
  geom_line(aes(color = antiviral), size = 0.75,alpha = 0.8) +

  # Vertical lines
  geom_vline(aes(xintercept = t_treat_since_inf), color = "green", linetype = "dotted", size = 1.2) +
  geom_vline(aes(xintercept = t_treat_since_inf + 5), color = "green", linetype = "dotted", size = 1.2) +

  # Facet by t_treat
  facet_wrap(~ t_treat, scales = "fixed") +

  scale_color_manual(values = c("1" = "#d62728", "0" = "#1f77b4"),  # treated = red, untreated = blue
                     labels = c("0" = "Untreated", "1" = "Treated")) +
  
  # Labels and theme
  labs(
    x = "Days after Infection",
    y = "Drug Amount in Plasma (mg)",
    color = "Antiviral"#, fill = "Antiviral"
  ) +
 theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='none',
        #axis.title.y = element_text(size=11,family="sans"),
       # axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  )+ theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 14),        # X-axis label font size
  axis.title.y = element_text(size = 14),        # Y-axis label font size
  axis.text = element_text(size = 14) ,  
  strip.text = element_text(size = 14),# Axis tick labels font size
  legend.title = element_text(size = 14),  # Legend title font size
  legend.text = element_text(size = 12),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
)  + xlim(0,30)

plot_pop_drug
```

###########################################################
###################### Individual Plots ###################
###########################################################


######### Read in the individual parameters
```{r}
Est <- read.csv("estimatedIndividualParameters.txt")
#Original <- read.csv("Paxlovid_match_rpart.csv", row.names = 1)
Sim <- read.csv("simulatedIndividualParameters.txt")
```


############### Function ###################
```{r}
Intfun_ind_pkpd <-function(pars){
  

  beta <- 10^as.numeric(pars[1])#MODE,
  pi <- 10^as.numeric(pars[2])
  phi <- 10^as.numeric(pars[3])
  rho <- 10^as.numeric(pars[4])
  delta <- as.numeric(pars[5])
  tau <- as.numeric(pars[6])
  prf <- as.numeric(pars[7])
  
  Antiviral <- as.numeric(pars[8])
  t_treat <-as.numeric(pars[9])
  
  
  # Fixed PKPD parameters from the paper
   
  E_max <- 0.999
  K_e <- 9.98
  K_PL <- 1.58
  K_LP <- 1.22
  K_CL <- 4.96
  Vol <- 41743
  MolMass <- 499.5
  n <- 3.16
  IC50 <- 0.083  # in µM (43.6 nM converted)
  Dose <- 300     # in mg
  
  # Time settings
  t_treat <- round(t_treat, digit =1 )
  tau <- round(tau, digit = 1)
  rec <- t_treat + tau	  #
  Tmin <- 0
  Tmax <- 40
  step_size <- 0.1
  times <- seq(Tmin, Tmax, step_size)
  


  
  treatment <- data.frame(times = times, h = rep(0, length(times)))
  #treatment$h <- ifelse(times < rec, 0, 1)
  treatment$h <- ifelse(times < rec | times > rec + 5, 0, 1) * Antiviral #5day treatment
  
  
  h_t2 <- approxfun(treatment$times, treatment$h, rule = 2)
  
  dose_times <- seq(rec, rec + 4.5, by = 0.5)  # t_treat, t_treat+0.5, ..., t_treat+4.5
  dose_events <- data.frame(var = "A_G", time = dose_times, value = Dose, method = "add")

 derivs <- function(times, y, pars, h_t2) {
    with(as.list(c(pars, y)), {
      # Treatment indicator
      h2 <- h_t2(times)
      
      # PK equations
      dA_G <- -K_e * A_G
      dA_P <- K_e * A_G + K_LP * A_L - (K_CL + K_PL) * A_P
      dA_L <- K_PL * A_P - K_LP * A_L
      
      # Compute plasma concentration
      C_P_mg_per_mL <- A_P / Vol
      C_P_uM <- C_P_mg_per_mL * (1e6 / MolMass)
      
      # Compute efficacy (eps)
      eps <- E_max * (C_P_uM^n) / ((prf * IC50)^n + C_P_uM^n)
      
      # Apply treatment window (eps_t)
      eps_t <- ifelse(h2 == 1, eps, 0)
      
      # Viral dynamics equations
      dT <- -beta * T * V - phi * I * T + rho * R
      dR <- phi * I * T - rho * R
      dE <- beta * T * V - 4 * E
      dI <- 4 * E - delta * I
      dV <- (1 - eps_t) * pi * I - 15 * V
      
      return(list(c(dT, dR, dE, dI, dV, dA_G, dA_P, dA_L), eps_t = eps_t))
    })
  }
  
  # Initial conditions
  y <- c(T = 10^7, R = 0, E = 0, I = 0, V = 1, A_G = 0, A_P = 0, A_L = 0)
  
 out <- ode(y = y, times = round(times,digit=1), func = derivs, parms = pars, h_t2 = h_t2, events = list(data = dose_events), method = "bdf")
  
  out_df <- as.data.frame(cbind(time = out[,1], aV1 = log10(out[,6]), eps_t = out[,"eps_t"]))
  
  return(out_df)
}

```
############ Calculate individual efficacy ###############
```{r}
#Calculate the efficacy for everyone
# Create a sequence of prf values from 5 to 100 in steps of 5
Treated <- mode_pars %>% filter(Antiviral == 1)
ids <- Treated$ID 

# Create empty df to store the result
ind_eff <- data.frame(id = ids, E_ave = NA_real_)

pb <- txtProgressBar(min = 0, max = length(ids), style = 3)

for (i in 1:length(ids)) {
  current_id <- ids[i]
  parameter <- mode_pars %>% filter(ID == current_id)
  
   t_treat <- round(parameter[9], digit = 1)+ round(parameter[6], digit = 1)
  result <- Intfun_ind_pkpd(parameter)
  
  E_ave <- ComputeEave(result, t_treat$t_treat)
  
  ind_eff$E_ave[i] <- E_ave
  setTxtProgressBar(pb, i)
}
close(pb)
print(ind_eff)
```

########### Select subgroups of parameters for fittings #########
```{r}
selected_sim <- Sim %>%
  dplyr::select(log10beta,log10pi,log10phi,log10rho,delta,tau,prf,antiviral,T_treat,id) %>%
  rename(ID=id)
selected_sim
mode_pars <- data.frame(log10beta = Est$log10beta_mode,
  log10pi= Est$log10pi_mode,
  log10phi=Est$log10phi_mode,
  log10rho=Est$log10rho_mode,
  delta_mode=Est$delta_mode,
  tau_mode=Est$tau_mode,
  prf_mode=Est$prf_mode,
  Antiviral = Est$antiviral,
  t_treat = Est$T_treat,
  ID = Est$id
)
mode_pars
```

############ Individual Fits: plotting function ######################

```{r}

ind_fit_plt1<-function(Estimated, Simulated, original){
  
  Fit <- list()
  for(i in 1:nrow(Estimated)){

    pars <- Estimated[i,] #Est should be a mode parameter dataframe with correct orders of paramters
    
    fitted <- Intfun_ind_pkpd(par
    index <- which(fitted$aV1 < -2)[1]
    # Check if an index was found, and then replace the value
    if (!is.na(index)) {
      fitted$aV1[index] <- -2
    } 
    
    d1 <- fitted
    
    
    Code <- Estimated$ID[i]
    original_subset <- original %>%
      filter(hosp_id == Estimated$ID[i]) %>%
      mutate(Day_since_inf = Day_since_detection + round(as.numeric(pars["tau_mode"]) ), digit = 1)
    
    S <- 50 ###50 repeats

    P <- matrix(NA,nrow=Tmax*(1/step_size)+1,ncol=S) #pre-sym and after-sym
    
    for(j in 1:S){
      pars <- c(beta=10^Simulated$log10beta[j+S*(i-1)],
                phi=10^Simulated$log10phi[j+S*(i-1)],
                rho=10^Simulated$log10rho[j+S*(i-1)],
                delta=Simulated$delta[j+S*(i-1)],
                pi=10^Simulated$log10pi[j+S*(i-1)],
                tau=Simulated$tau[j+S*(i-1)],
                prf=Simulated$prf[j+S*(i-1)],
                
                Antiviral=Simulated$antiviral[j+S*(i-1)],
                t_treat=Simulated$T_treat[j+S*(i-1)])
      
      out  <- Intfun_ind_pkpd(pars)
      #out_pre  <- Covfun_pre(pars)
      P[,j] <- out$aV1
    }
    
    Min95  <- apply(P,1,function(x){quantile(x,0.005,na.rm = TRUE)})
    Min90  <- apply(P,1,function(x){quantile(x,0.05,na.rm = TRUE)})
    Max90  <- apply(P,1,function(x){quantile(x,0.95,na.rm = TRUE)})
    Max95  <- apply(P,1,function(x){quantile(x,0.995,na.rm = TRUE)})
    Mean50 <- apply(P,1,function(x){quantile(x,0.50,na.rm=T)})
    fit <- cbind(d1,Min95,Min90,Max90,Max95,Mean50,Code)
    fit$Min95[fit$Min95 < -2] <- -2
    fit$Min90[fit$Min90 < -2] <- -2
    fit <- merge(fit, original_subset[,c("hosp_id","Day_since_inf","VL")], by.x=c("Code", "time"), by.y = c("hosp_id", "Day_since_inf"),all=TRUE)
    
    Fit[[i]] <- data.frame(fit)
    
    
  }
  
  ind_fit <- map_df(Fit, ~as.data.frame(.x))
  
  ind_fit$Code <- as.factor(ind_fit$Code)
  return(ind_fit)
}
```
                              
```{r}
    Tmin <- 0
    Tmax <- 40
    step_size <- 0.1
    times<-c(seq(Tmin,Tmax,step_size))
individual_fit <- ind_fit_plt1(mode_pars,selected_sim,Original)
```

# The individual Trajecotries data in "individual_fit" was used to create sup-Fig-2 and sup-Fig-3


```{r}
 Ori <- individual_fit %>%
  filter(if_all(VL, ~ !is.na(.))) %>%
  dplyr::select(Code, time, VL)

                              
individual_fit_selected <- individual_fit %>%
  filter(if_all(c(aV1,Min95,Max95), ~ !is.na(.))) %>%
  dplyr::select(Code, time, aV1, Min95, Max95, Mean50)

                              
Treat_info <- mode_pars %>%
  mutate(t_treat_after_inf = ifelse(Antiviral==1,(round(tau_mode,digit=1)+t_treat),NA)) %>%
  dplyr::select(ID,t_treat_after_inf) %>%
  rename(Code = ID) %>%
  filter(!is.na(t_treat_after_inf)) 
Treat_info$Code <- as.factor(Treat_info$Code)

 # Sub data for treated group                             
individual_fit_selected_1 <- individual_fit %>% 
  filter(Code %in% Treat_info$Code) %>%
  ungroup()
individual_fit_selected_1<- merge(individual_fit_selected_1, Treat_info, by = "Code")   
individual_fit_selected_1$num <- as.numeric(individual_fit_selected_1$Code)
individual_fit_selected_1$num <- as.factor(individual_fit_selected_1$num)

# Sub data for untreated group
individual_fit_selected_2 <- individual_fit %>% 
  filter(!Code %in% Treat_info$Code) 
individual_fit_selected_2$num <- as.factor(as.numeric(individual_fit_selected_2$Code))                           
```    
###################################################
################ Plot #############################                             
###################################################

```{r}
pdf(paste0("Figure/new/Individual Fit_new", ".pdf"), 16, 16)
for (i in seq(1, length(unique(individual_fit_selected_1$num)), 120)) {
  print(
    ggplot(individual_fit_selected_1[individual_fit_selected_1$num %in% levels(individual_fit_selected_1$num)[i:(i+119)],], aes(x=time)) +
      
  # Add shaded area for Min/Max range
  geom_ribbon(aes( ymin = Min95, ymax = Max95,fill = "95% CI"), alpha = 0.7 ) +
  # Add the fit line
  geom_line(aes( y = aV1, color = "Best Fit"), size = 0.3, alpha = 0.8) +
  geom_hline( linetype = "dotted", aes(yintercept = 1.31, color = "Detection Limit"), size = 0.6, alpha = 1)+
  
  # Add individual's point
  geom_point( aes( y = VL,color = "Original Data"), 
             shape = 16, size = 1, alpha = 2) +
  geom_vline(aes(xintercept = t_treat_after_inf, color = "Treatment Date"),linetype = "dotted", size = 0.5) +
 
  facet_wrap(~num, ncol=12, nrow=20) +
  
  labs(
    title = "Individual Fit with Shaded Uncertainty",
    x = "Days after Infection",
    y = "Viral Load",
    fill = "",
    color = ""
  ) +
 theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='right',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() ) +
  
  scale_fill_manual( values = c("95% CI" = "pink")) +
  
  scale_color_manual(
   
    values = c("Best Fit" = "red", "Original Data" = "blue", "Treatment Date" = "green", "Detection Limit" = "grey")
  ) +
  ylim(-2,10) + xlim(0,30)
      
  )
}
dev.off()
```                              
```{r}
pdf(paste0("Figure/new/Individual Fit_untreated_new", ".pdf"), 16, 16)
for (i in seq(1, length(unique(individual_fit_selected_2$num)), 120)) {
  print(
    ggplot(individual_fit_selected_2[individual_fit_selected_2$num %in% levels(individual_fit_selected_2$num)[i:(i+119)],], aes(x=time)) +
      
  # Add shaded area for Min/Max range
  geom_ribbon(aes( ymin = Min95, ymax = Max95,fill = "95% CI"), alpha = 0.7 ) +
  # Add the fit line
  geom_line(aes( y = aV1, color = "Best Fit"), size = 0.3, alpha = 0.8) +
  geom_hline( linetype = "dotted", aes(yintercept = 1.31, color = "Detection Limit"), size = 0.6, alpha = 1)+
  
  # Add individual's point
  geom_point( aes( y = VL,color = "Original Data"), 
             shape = 16, size = 1, alpha = 2) +
 # geom_vline(aes(xintercept = t_treat_after_inf, color = "Treatment Date"),linetype = "dotted", size = 0.5) +
 
  facet_wrap(~num, ncol=12, nrow=20) +
  
  labs(
    title = "Individual Fit with Shaded Uncertainty",
    x = "Days after Infection",
    y = "Viral Load",
    fill = "",
    color = ""
  ) +
 theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='right',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() ) +
  
  scale_fill_manual( values = c("95% CI" = "pink")) +
  
  scale_color_manual(
   
    values = c("Best Fit" = "red", "Original Data" = "blue", "Treatment Date" = "green", "Detection Limit" = "grey")
  ) +
  ylim(-2,10) + xlim(0,30)
      
      
  )
  
}
dev.off()
```    


#############################################
################## S-Figure-5 ###############
#############################################
```{r}
estimate_raw <- read.csv("estimatedIndividualParameters.txt")
estimate_raw$tau_mode <-  round(mean(estimate_raw$tau_mode),digit = 1)
estimate_raw$T_treat[estimate_raw$antiviral == 1] <- round(mean(estimate_raw$T_treat[estimate_raw$antiviral == 1]), digits = 1)
```

####################################################################
#### Function to compute CI based on a subgroup of data ############
####################################################################
```{r}
ind_fit_plt2<-function(Estimated){
  
  Fit <- list()

    S <- nrow(Estimated)

    P <- matrix(NA,nrow=Tmax*(1/step_size)+1,ncol=S) #pre-sym and after-sym
    #P <- matrix(NA,nrow=(Tmax+1),ncol=S)
    for(j in 1:S){
      pars <- Estimated[j,] #Est should be a mode parameter dataframe with correct orders of paramters
      
      out  <- Intfun_ind_pkpd(pars)
      #out_pre  <- Covfun_pre(pars)
      P[,j] <- out$aV1
    }
    
    Min95  <- apply(P,1,function(x){quantile(x,0.005,na.rm = TRUE)})
    Min90  <- apply(P,1,function(x){quantile(x,0.25,na.rm = TRUE)})
    Max90  <- apply(P,1,function(x){quantile(x,0.75,na.rm = TRUE)})
    Max95  <- apply(P,1,function(x){quantile(x,0.995,na.rm = TRUE)})
    Mean50 <- apply(P,1,function(x){quantile(x,0.50,na.rm=T)})
    fit <- cbind(times,Min95,Min90,Max90,Max95,Mean50)
    
    Fit <- data.frame(fit)
    Fit$Min95[Fit$Min95 < -2] <- -2
    Fit$Min90[Fit$Min90 < -2] <- -2
    Fit$Group <- Estimated$antiviral[1]
    Fit$treatment <- ifelse(Fit$Group == 1, "treated", "untreated")
  return(Fit) }
```
##########################################################
############### Stratification by vac status##############
##########################################################
```{r}
full_ind <- subset(estimate_raw, vaccination == c("Full"))
full_ind_t <- subset(full_ind, antiviral == 1)
full_ind_t
Booster_ind <- subset(estimate_raw, vaccination == c("Booster"))
Booster_ind_t <- subset(Booster_ind, antiviral == 1)
Booster_ind_t
Not_Full_ind <- subset(estimate_raw, vaccination == c("Notfull"))
Not_Full_ind_t <- subset(Not_Full_ind, antiviral == 1)
Not_Full_ind_t
Full_Booster_ind <- subset(estimate_raw, vaccination %in% c("Booster", "Full"))
Full_Booster_ind_t <- subset(Full_Booster_ind, antiviral == 1)
Full_Booster_ind_t
```

```{r}
Full_treated <- ind_fit_plt2(full_ind_t[,c(16:22,38,36)])
Not_Full_treated <- ind_fit_plt2(Not_Full_ind_t[,c(16:22,38,36)])
Booster_treated <- ind_fit_plt2(Booster_ind_t[,c(16:22,38,36)])
Full_Booster_treated <- ind_fit_plt2(Full_Booster_ind_t[,c(16:22,38,36)])
```

```{r}
Full_treated$vaccination <- "Fully vaccinated"
Not_Full_treated$vaccination <- "Not fully vaccinated"
Booster_treated$vaccination <- "Booster shot"
combined_vac <- rbind(Full_treated,Not_Full_treated,Booster_treated)
combined_vac

Full_Booster_treated$vaccination <- "Full vaccinated or Boosted"
combined_vac_2 <- rbind(Full_Booster_treated,Not_Full_treated)
combined_vac_2
```

```{r}
# Create the plot
vac_plot <- ggplot(combined_vac_2, aes(x = times, y = Mean50, fill = vaccination)) +
  
  geom_ribbon(aes(ymin = Min95, ymax = Max95, fill = vaccination), alpha = 0.12) +
 # geom_ribbon(aes(ymin = Min90, ymax = Max90, fill = treatment), alpha = 0.17) +
  geom_line(aes(color = vaccination),alpha = 0.8, linewidth = 0.75) +
  #geom_vline(aes(xintercept = t_treat_since_inf,  color = t_treat),linetype = "dotted", linewidth = 1.2) +
  geom_vline(aes(xintercept = 5.2), color = "green",linetype = "dotted", linewidth = 1) +
  geom_vline(aes(xintercept = 10.2), color = "green",linetype = "dotted", linewidth = 1) +
  geom_hline(aes(yintercept = 1.31),color = "grey",linetype = "dotted", size = 1) +
  labs(
    #title = "Visualization of Viral Load Trajectory",
    x = "Days after Infection",
    y = "log10 Viral Load (RNA copies/ml)",
    
  ) +
  theme_minimal() +
   theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='bottom',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  ) +
  theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 14),        # X-axis label font size
  axis.title.y = element_text(size = 14),        # Y-axis label font size
  axis.text = element_text(size = 14) ,           # Axis tick labels font size
  legend.title = element_blank(),  # Legend title font size
  legend.text = element_text(size = 12),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
) + ylim(-2,11) + xlim(0,30)
vac_plot
```
##########################################################
############### Stratification by age ####################
##########################################################
```{r}
old_ind <- subset(estimate_raw, age_cat == ">=65")
old_ind_t <- subset(old_ind, antiviral == 1)
hist(old_ind$antiviral)
old_ind_t
old_ind_ut <- subset(old_ind, antiviral == 0)
old_ind_ut
young_ind <- subset(estimate_raw, age_cat == "18-64")
young_ind_t <- subset(young_ind, antiviral == 1)
hist(young_ind$antiviral)
young_ind_t
young_ind_ut <- subset(young_ind, antiviral == 0)
young_ind_ut
```
```{r}
old_untreated <- ind_fit_plt2(old_ind_ut[,c(16:22,38,36)])
old_treated <- ind_fit_plt2(old_ind_t[,c(16:22,38,36)])
young_untreated <- ind_fit_plt2(young_ind_ut[,c(16:22,38,36)])
young_treated <- ind_fit_plt2(young_ind_t[,c(16:22,38,36)])
```
```{r}
old_treated$age <- ">=65"
old_untreated$age <- ">=65"
young_untreated$age <- "18-64"
young_treated$age <- "18-64"
```
```{r}
old_combined <- rbind(old_untreated,old_treated)
young_combined <- rbind(young_untreated,young_treated)

age_treat <- rbind(old_treated,young_treated)
age_untreat <- rbind(old_untreated,young_untreated)
```

```{r}
age_plot <- ggplot(age_treat, aes(x = times, y = Mean50, fill = age)) +
  
  geom_ribbon(aes(ymin = Min95, ymax = Max95, fill = age), alpha = 0.12) +
 # geom_ribbon(aes(ymin = Min90, ymax = Max90, fill = treatment), alpha = 0.17) +
  geom_line(aes(color = age),alpha = 0.8, linewidth = 0.75) +
  #geom_vline(aes(xintercept = t_treat_since_inf,  color = t_treat),linetype = "dotted", linewidth = 1.2) +
  #geom_vline(aes(xintercept = t_treat_since_inf,  color = t_treat),linetype = "dotted", linewidth = 1.2) +
  geom_vline(aes(xintercept = 5.2), color = "green",linetype = "dotted", linewidth = 1) +
  geom_vline(aes(xintercept = 10.2), color = "green",linetype = "dotted", linewidth = 1) +
  geom_hline(aes(yintercept = 1.31),color = "grey",linetype = "dotted", size = 1) +
  labs(
    #title = "Visualization of Viral Load Trajectory",
    x = "Days after Infection",
    y = "log10 Viral Load (RNA copies/ml)",
    
  ) +
  theme_minimal() +
   theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='bottom',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  ) +
  theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 14),        # X-axis label font size
  axis.title.y = element_text(size = 14),        # Y-axis label font size
  axis.text = element_text(size = 14) ,           # Axis tick labels font size
  legend.title = element_text(size = 14),  # Legend title font size
  legend.text = element_text(size = 12),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
) + ylim(-2,11) + xlim(0,30)
age_plot
```
