##### This code perform regressions of treatment efficacy (as result of TREIV-PKPD Fitting) on Patients' characteristcs ####
##### Also perform the regression of model parameters on Patients' characteristcs #####
##### Fig 4, S-Fig-4, S-Fig-5 ######

```{r}
library(dplyr)
library(ggplot2)
library(GGally)
library(tidyr)
library(finalfit)
library(emmeans)
```

```{r}
estimate <- read.csv("C:/Users/YIYU.LIAO/OneDrive - Nanyang Technological University/Desktop/Shanghai data/run/rpart_pkpd/prf_nocov_r3_new_correctdose/IndividualParameters/estimatedIndividualParameters.txt")
estimate <- estimate %>% filter(antiviral == 1)
```

# read in individual efficacy which was calculated in previous section
```{r}
ind_eff <- read.csv("ind_eff.csv")
ind_eff
estimate_merge <- merge(estimate, ind_eff, by = "id")
estimate_merge
```

######## Plot S-Fig-4 ###########
```{r}
numeric_est <- estimate_merge %>% dplyr::select(log10beta_mode,log10pi_mode,log10phi_mode,log10rho_mode,delta_mode,tau_mode,  prf_mode, E_ave,T_treat,
                                                age,comorbidity_sum,vax_doses) %>% rename(
                                                  beta = log10beta_mode,
                                                  pi = log10pi_mode,
                                                  phi = log10phi_mode,
                                                  rho = log10rho_mode,
                                                  delta = delta_mode,
                                                 # tau = tau_mode,
                                                  prf = prf_mode
                                                )
```
```{r}
numeric_est_long <- numeric_est  %>% dplyr::select(beta,pi,phi,rho,delta,prf) %>% rename(log10beta = beta, log10pi = pi, log10rho=rho,log10phi = phi)%>% pivot_longer(cols = everything(), names_to = "parameter", values_to = "value") 

numeric_est_long$parameter <- factor(numeric_est_long$parameter, levels = c("log10beta", "log10pi", "log10rho","log10phi","delta","prf" ))




hist_pop <- ggplot(numeric_est_long, aes(x = value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~ parameter, scales = "free") +
   coord_cartesian(ylim = c(0, 60)) +    
  theme_minimal() +
  labs( x = NULL, y = "Count")  +
  theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='none',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  ) +
 theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 12),        # X-axis label font size
  axis.title.y = element_text(size = 12),        # Y-axis label font size
  axis.text = element_text(size = 12) ,           # Axis tick labels font size
 strip.text = element_blank(),
  #strip.background = element_rect(fill = "lightblue", colour = "black"),
  plot.margin = margin(t = 1, r = 1, b = 1, l = 1, unit = "cm")
) + theme(panel.spacing = unit(1.5, "cm"))
hist_pop
```

######################################################################################################################
############ Final-fit to conduct regressions of model parameters and efficacy on patients' characteristics ##########
######################################################################################################################

# Standardize the order of characteristics
```{r}
estimate_merge$age_cat <- factor(estimate_merge$age_cat, levels = c("18-64", ">=65"))
estimate_merge$comorbidity_cat <- factor(estimate_merge$comorbidity_cat, levels = c("0", ">=1"))
estimate_merge$vaccination <- factor(estimate_merge$vaccination, levels = c("Notfull", "Full", "Booster"))
```

####### Final fit ############
```{r}
explanatory = c(  "age_cat" ,"sex","comorbidity_cat",  "vaccination")
dependent = 'log10beta_mode'
estimate_merge %>%
  finalfit(dependent, explanatory, metrics=TRUE) 
```
```{r}
explanatory = c(  "age_cat" ,"sex","comorbidity_cat",  "vaccination")
dependent = 'log10pi_mode'
estimate_merge %>%
  finalfit(dependent, explanatory, metrics=TRUE) 
```
```{r}
explanatory = c(  "age_cat" ,"sex","comorbidity_cat",  "vaccination")
dependent = 'log10rho_mode'
estimate_merge %>%
  finalfit(dependent, explanatory, metrics=TRUE) 
```
```{r}
explanatory = c(  "age_cat" ,"sex","comorbidity_cat",  "vaccination")
dependent = 'log10phi_mode'
estimate_merge %>%
  finalfit(dependent, explanatory, metrics=TRUE) 
```
```{r}
explanatory = c(  "age_cat" ,"sex","comorbidity_cat",  "vaccination")
dependent = 'delta_mode'
estimate_merge %>%
  finalfit(dependent, explanatory, metrics=TRUE) 
```
```{r}
explanatory = c(  "age_cat" ,"sex","comorbidity_cat",  "vaccination")
dependent = 'tau_mode'
estimate_merge %>%
  finalfit(dependent, explanatory, metrics=TRUE) 
```
```{r}
explanatory = c(  "age_cat" ,"sex","comorbidity_cat",  "vaccination")
dependent = 'prf_mode'
estimate_merge %>%
  finalfit(dependent, explanatory, metrics=TRUE) 
```
```{r}
explanatory = c(  "age_cat" ,"sex","comorbidity_cat",  "vaccination")
dependent = 'E_ave'
estimate_merge %>%
  finalfit(dependent, explanatory, metrics=TRUE) 
```

######## Redo the multivariate regression of efficacy and plot the boxplot ############
############################
######## Figure 4 ##########
############################

######### 4-A: Vaccination status ##########
```{r}
estimate_merge$vaccination <- factor(estimate_merge$vaccination, levels = c("Notfull", "Full", "Booster"))

linear_vac_cat <- lm(E_ave ~ vaccination, data = estimate_merge)
emm <- emmeans(linear_age_cat, ~ vaccination)
emm_df <- as.data.frame(emm)

estimate_merge$vaccination <- factor(estimate_merge$vaccination, levels = c("Notfull", "Full", "Booster"))

vax_cat_plot <- ggplot(estimate_merge) +
  geom_boxplot(aes(x = vaccination, y = E_ave),outlier.shape = NA, fill = "grey90", color = "grey40") +
  geom_jitter(aes(x = vaccination, y = E_ave),width = 0.1, alpha = 0.2, color = "darkred") +
  geom_point(data = emm_df, aes(x = vaccination, y = emmean, group = 1), 
             color = "blue", size = 3) +
  geom_errorbar(data = emm_df,
                aes(x = vaccination, ymin = lower.CL, ymax = upper.CL, group = 1),
                width = 0.3, color = "blue", size = 1) +
  geom_line(data = emm_df, aes(x = vaccination, y = emmean, group = 1), size = 0.7,
            color = "blue", linetype = "dashed") +
  labs(#title = "Efficacy by Age Group",
       x = "Vaccination Group",
       y = "Efficacy") +
  theme_minimal() +
   theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='bottom',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  ) +
  theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 14),        # X-axis label font size
  axis.title.y = element_text(size = 14),        # Y-axis label font size
  axis.text = element_text(size = 14) ,           # Axis tick labels font size
  legend.title = element_blank(),  # Legend title font size
  legend.text = element_text(size = 12),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
) 
vax_cat_plot
```
########## 4-B: Age ##########
```{r}
linear_age_cat <- lm(E_ave ~ age_cat +sex + vaccination +comorbidity_cat, data = estimate_merge)
summary(linear_age_cat)
# Get model-predicted means and CIs
emm <- emmeans(linear_age_cat, ~ age_cat)
emm_df <- as.data.frame(emm)

estimate_merge$age_cat <- factor(estimate_merge$age_cat, levels = c("18-64", ">=65"))

age_cat_plot <- ggplot(estimate_merge) +
  geom_boxplot(aes(x = age_cat, y = E_ave), outlier.shape = NA, fill = "grey90", color = "grey40") +
  geom_jitter(aes(x = age_cat, y = E_ave), width = 0.1, alpha = 0.2, color = "darkred") +
  geom_point(data = emm_df, aes(x = age_cat, y = emmean, group = 1), 
             color = "blue", size = 3) +
  geom_errorbar(data = emm_df,
                aes(x = age_cat, ymin = lower.CL, ymax = upper.CL, group = 1),
                width = 0.3, color = "blue", size = 1) +
  geom_line(data = emm_df, aes(x = age_cat, y = emmean, group = 1), size = 0.7,
            color = "blue", linetype = "dashed") +
  labs(#title = "Efficacy by Age Group",
       x = "Age Group",
       y = "Efficacy") +
  theme_minimal() +
   theme(axis.text = element_text(colour = "black"),
        axis.ticks = element_line(colour = "black"),
        axis.line = element_line(colour = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position='bottom',
        axis.title.y = element_text(size=11,family="sans"),
        axis.title.x = element_text(size=11,family="sans"),
  theme_minimal() 
  ) +
  theme(
 # plot.title = element_text(size = 16),          # Title font size
  axis.title.x = element_text(size = 14),        # X-axis label font size
  axis.title.y = element_text(size = 14),        # Y-axis label font size
  axis.text = element_text(size = 14) ,           # Axis tick labels font size
  legend.title = element_blank(),  # Legend title font size
  legend.text = element_text(size = 12),
  plot.margin = margin(t = 0.5, r = 0.5, b = 0.5, l = 0.5, unit = "cm")
) 
age_cat_plot
```

